Effects of Reynolds Number on the Energy Conversion and Near-Wake Dynamics of a High Solidity Vertical-Axis Cross-Flow Turbine

### 2015-12-14 paper.tex
Shorten intro section on Re dependence; resolves #47

\subsection{Reynolds Number Effects}

In this study we are using a simple cross-flow turbine constructed from
hydrofoils, i.e., the rotor is lift-driven in contrast to drag-driven. A logical
starting point for understanding the effects of $Re$ on overall turbine loading
is to first consider simplified cases of individual foils, and their static and
dynamic loading.

\subsubsection{On Static Foil Behavior}

Hydro- or airfoil section performance is generally characterized by the
lift-to-drag force ratio $L/D$ (or the ratio of lift and drag coefficients,
$C_l/C_d$), which is primarily a function of the foil's angle of attack
$\alpha$. The angles of attack for cross-flow turbine blades are a function of
turbine tip speed ratio, induction (slowing/turning of the free stream flow
prior to reaching the turbine), and turbine rotation or azimuthal angle
$\theta$. Typically, static airfoil data shows that the static stall
angle---the angle just beyond that at which $L/D$ is maximum---increases with
blade chord Reynolds number $Re_c$ \cite{Jacobs1937}.

A review of Reynolds number effects on airfoil behavior is presented in
\cite{Lissaman1983}. In general, airfoil performance is enhanced as the boundary
layer on the foil transitions to turbulence closer to the leading edge, which
enables it to advance further downstream against the adverse pressure gradient
on the suction side, delaying separation to higher angles of attack. For smooth
airfoils, this transition can cause a dramatic increase in foil performance, at
a blade chord Reynolds number on the order of $10^5$ \cite{McMasters1980}, shown
in Figure~\ref{fig:McMasters}. Note that there is a distinct lack of highly
reliable data for airfoils in this transitional regime and below. An evaluation
of the various databases relevant to cross-flow turbines is presented in
\cite{Bedon2014}.

@@

\subsubsection{On Dynamic Foil Behavior}

Static foil performance does not tell the whole story for a cross-flow turbine.
The azimuthal, and therefore temporal variation of $\alpha$ in a cross-flow
turbine implies dynamic loading. Furthermore, since the range of angle of attack
encountered by a rotating blade increases with decreasing tip speed ratio, the
blades commonly undergo dynamic stall for tip speed ratios at and below those of
maximum rotor torque \cite{Para2002}. We then must understand the effects of
Reynolds number on the dynamic stall process.

Bousman \cite{Bousman2000-evaluation} states that dynamic stall is relatively
insensitive to Reynolds number for $Re=1.0 \times 10^5$--$2.5 \times
10^5$---judging from measurements on a pitching VR-7 foil in a wind and water
tunnel---since the loading is dominated by vortex shedding. However, Singleton
and Yeager Jr. \cite{Singleton2000} state that Reynolds number effects on
dynamic stall remains an unsolved question.

\subsubsection{On Wake Flows}

@@

Chamorro et al. \cite{Chamorro2012} showed that turbulence statistics in an AFT
wake became $Re$-independent at $Re_D \approx 1 \times 10^5$, and that mean
velocity profiles became $Re$-independent slightly earlier at $Re_D \approx 5
\times 10^4$. Note that in this study the tip chord Reynolds number was not
reported, but is estimated to be $Re_c \approx 4 \times 10^4$ at $Re_D=1 \times
10^5$.

>>>>

\subsection{Modes of Reynolds Number Dependence}

We are looking at how Reynolds number scaling affects both the blade loading,
i.e., turbine performance, and the near-wake. Typically, static airfoil data
shows that the static stall angle---the angle just beyond that at which the
lift-to-drag ratio is maximum---increases with blade chord Reynolds number
$Re_c$ \cite{Jacobs1937}. A review of Reynolds number effects on airfoil
behavior is presented in \cite{Lissaman1983}. In general, airfoil performance is
enhanced as the boundary layer on the foil transitions to turbulence closer to
the leading edge, which enables it to advance further downstream against the
adverse pressure gradient on the suction side, delaying separation to higher
angles of attack. For smooth airfoils, this transition can cause a dramatic
increase in foil performance, at a blade chord Reynolds number on the order of
$10^5$ \cite{McMasters1980}, shown in Figure~\ref{fig:McMasters}. Note that
there is a distinct lack of highly reliable data for airfoils in this
transitional regime and below. An evaluation of the various databases relevant
to cross-flow turbines is presented in \cite{Bedon2014}.

@@

Static foil performance does not tell the whole story for a cross-flow turbine.
The azimuthal, and therefore temporal variation of $\alpha$ in a cross-flow
turbine implies dynamic loading, encountering dynamic stall for tip speed ratios
near and below those of maximum rotor torque \cite{Para2002}. Bousman
\cite{Bousman2000-evaluation} states that dynamic stall is relatively
insensitive to Reynolds number for $Re=1.0 \times 10^5$--$2.5 \times
10^5$---judging from measurements on a pitching VR-7 foil in a wind and water
tunnel---since the loading is dominated by vortex shedding. However, Singleton
and Yeager Jr. \cite{Singleton2000} state that Reynolds number effects on
dynamic stall remains an unsolved question.

@@

Chamorro et al. \cite{Chamorro2012} showed that turbulence statistics in an AFT
wake became $Re$-independent at $Re_D \approx 1 \times 10^5$, and that mean
velocity profiles became $Re$-independent slightly earlier at $Re_D \approx 5
\times 10^4$. Note that in this study the tip chord Reynolds number was not
reported, but is estimated to be $Re_c \approx 4 \times 10^4$ at $Re_D=1 \times
10^5$. It is therefore our objective to observe similar scaling relationships
for a cross-flow turbine near-wake.
